---
title: "First R notebook for Ames housing dataset"
output:
  html_document:
    df_print: paged
---



```{r}
#Data preprocessing
#importing the libraries 
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(randomForest)
library(xgboost)
```
```{r}
#importing the train and test datasets
train <- read.csv("train.csv", stringsAsFactors = F)
test <- read.csv("test.csv", stringsAsFactors = F)
head(train)

```
```{r}
dim(train)
```
```{r}
str(train)
```
```{r}
testids <- test$Id
test$Id <- NULL
train$Id <- NULL
test$SalePrice <- NA
ncol(test)
ncol(train)
```
```{r}
#Combining the train and test sets
combined <- rbind(train, test) 
```
```{r}
#examining the distribution of dependent variable
#p <- hist(combined$SalePrice, main = "Histogram of Sale Prices", xlab = "SalePrice")
ggplot(data = combined, aes(x = SalePrice)) + geom_histogram(binwidth = 10000) + scale_x_continuous(breaks = seq(0,800000, by=100000))

```
```{r}
numeric <- which(sapply(combined, is.numeric))
numericvars <- combined[,numeric]
numericvars
correlation_numericvars <- cor(numericvars, use = "pairwise.complete.obs")
```
```{r}
correlation_sorted <- as.matrix(sort(correlation_numericvars[, 'SalePrice'], decreasing = TRUE))
highcorrelation <- names(which(apply(correlation_sorted, 1, function(x) abs(x)>0.5)))
correlation_numericvars <- correlation_numericvars[highcorrelation, highcorrelation]
```
```{r}
#getting a heatmap
corrplot.mixed(correlation_numericvars, tl.col='Black', tl.pos = 'lt')
```

#The variables which have maximum correlation with SalePrice(dependent variable) are OverallQual and GrLivArea
#Now we should examine the variation of Sale Price with these two variables
```{r}
ggplot(data=combined[!is.na(combined$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot() + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000))

```
*As expected there is positive correlation between Overall quality and Sale price. Also, there is an upward curve for the same.*

```{r}
ggplot(data=combined[!is.na(combined$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000))
```
*there seems to be two observations with very large value for living area but less sale price. Can we think of that as an outlier?*
*to determine if these two observations can be omitted, we need to look at other factors and to do so we have to find out which observations they are*

```{r}
library(ggrepel)
ggplot(data=combined[!is.na(combined$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000)) +
        geom_text_repel(aes(label = ifelse(combined$GrLivArea[!is.na(combined$SalePrice)]>4500, rownames(all), '')))
```

